Efficient and Scalable Entity Resolution with Fast Beta Linkage (fabl)

### 2021-06-15 methods_paper/fabl-paper.tex
nltcs and fabl paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{hyperref}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\setstretch{1}
\begin{flushright} 
    \end{flushright}
    \begin{center} \textbf{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}

    Brian Kundinger, Rebecca Steorts, and Jerome Reiter

    \end{center}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Entity resolution (or record linkage) is the task of identifying
duplicate records across multiple data sources. This task is
straightforward when provided unique identifiers or highly reliable
information, but becomes difficult when data sources are plagued with
error. Additionally, many entity resolution tasks are large in nature,
and many existing methods are too computational intensive to be used in
practice. In this paper, we propose a Bayesian model for record linkage
that uses hashing techniques and efficient storage mechanisms, and is
amenable to parallelization, making it scalable to large linkage tasks.

The proposed method is inspired by the \emph{Beta Record Linkage}
(\texttt{BRL}) method proposed in Mauricio Sadinle's 2017 ``Bayesian
Estimation of Bipartite Matching for Record Linkage''. To recognize this
lineage, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable''). In what follows, Section 2 describes previous
work in record linkage that lays the foundation for our method, Section
3 provides our model and the Gibbs sampler used for posterior inference,
and Section 4 describes strategies used for more efficient and scalable
implementation. Then, Section 5 demonstrates the accuracy and speed of
out method through two simulation studies, and Section 6 reveals
interesting properties of the method through a case study of El
Salvadoran casuality records. Lastly, our discussion highlights open
questions in the record linkage liturature and motivates future work.

\hypertarget{related-work}{%
\subsection{Related Work}\label{related-work}}

Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their approach was to transform the sets of records,
which often contain text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates an
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
score on the \(f^{th}\) field. For ease of modeling and computation, we
restrict these similarity scores to be discrete, ordinal variables, and
the construction of these is left to the modeler. We adopt the
convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
is common to use use binary 1-2 variables to indicate exact matching,
and 1-2-3 variables to provide an option for partial matching. For text
data, we calculate similarity based on Levenstein distance or some other
text similarity score, and bin these scores to integers for use in the
model. Jaro (1989) provided guidelines for appropriate binning for the
comparison vectors, but details of implementation are generally context
specific.

@@

One key strength of Sadinle's \texttt{BRL} is that it is garunteed to
produce a set of bipartite matchings that respect one-to-one
assumptions--without the need for any post-hoc procedures. Specifically,
in each iteration of his Gibbs sampler, he considers each record
\(j\in B\), removes from consideration the records \(i\in A\) that have
already been matched, and then samples a potential link. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

Outside of the Fellegi-Sunter framework, some researchers have developed
methods that model the data directly, not a derived set of comparison
vectors. As an early example, Steorts et al 2016 produced
\texttt{blink}, which had the advantage of a computational complexity
that grew linearly with the data (as opposed to quadratically within
Fellegi Sunter), but was plagued with slow mixing times and was unable
to incorporate text data. Later, Steorts et al 2020 improved on this
method with \texttt{dblink}, which was able to simulate text data within
its Gibbs sampler by drawing from an empirical distribution, and
remedied its slow mixing through a probabilistic blocking approach.
However, this method remains computationally intensive and requires
distributed computing to manage even moderate linkage problems in
acceptable time. Because these methods fall outside of the
Fellegi-Sunter framework, we do not consider them further in this paper.

\hypertarget{notation-and-assumptions}{%
\subsection{Notation and Assumptions}\label{notation-and-assumptions}}

@@

\hypertarget{model-specification}{%
\subsection{Model Specification}\label{model-specification}}

@@

\hypertarget{posterior-sampling}{%
\subsection{Posterior Sampling}\label{posterior-sampling}}

We work with the following factorization of the joint distribution:

\[p(\Gamma, \mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda) = p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]

@@

\hypertarget{bayes-estimate}{%
\subsection{Bayes Estimate}\label{bayes-estimate}}

@@

\hypertarget{efficient-computation}{%
\section{Efficient Computation}\label{efficient-computation}}

Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P = |\mathcal{H}|\) denote the total number of unique agreement
patterns. Note that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
that this bound does not scale with \(n_A\) or \(n_B\). Prior to
processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
and enumerate them \(h_1, \ldots, h_P\). When the \((i,j)\) pair
exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
Wherever possible, we conduct calculations over these \(P\) agreement
patterns rather than the \(n_A \times n_B\) record pairs.

@@

\hypertarget{gibbs-sampling}{%
\subsection{Gibbs Sampling}\label{gibbs-sampling}}

@@

\hypertarget{simulation-study-1---accuracy}{%
\section{Simulation Study 1 -
Accuracy}\label{simulation-study-1---accuracy}}

@@

\begin{figure}

{\centering \includegraphics[width=29.17in]{../notes/figures/sadinle_sim_plot}

\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:unnamed-chunk-1}
\end{figure}

\hypertarget{simulation-study-2---speed}{%
\section{Simulation Study 2 - Speed}\label{simulation-study-2---speed}}

@@

\begin{figure}

{\centering \includegraphics[width=17.75in]{../notes/figures/sadinle_speed_plot2}

@@

\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\begin{figure}

{\centering \includegraphics[width=29.17in]{../notes/figures/speed_plot_fixed_nB}

@@

\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\hypertarget{data-analysis}{%
\section{Data Analysis}\label{data-analysis}}

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualities, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casuality lists. To obtain a more accurate estimate of the
casualties then, we follow Sadinle's 2017 paper and link records of El
Salvadoran civilian casualties from two sources: El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

@@

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/m_posterior_smallP}

@@

\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:unnamed-chunk-4}
\end{figure}

To mirror the original implementation, we constructed the comparison
vectors using 4 levels of agreement for each field, according to the
thresholds provided in Figure XX. This took 422 seconds for our proposed
method and 240 for \texttt{BRL}. However, we observed that posterior
distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
parameters were nearly identical, and that of the
\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
realized in the data. This leads us to believe that such high number of
aggreement levels creates unnecesary distictions in the data and makes
the comparison vectors less interpretable. Therefore we re-ran our
analysis with fewer agreement levels for each field (see Figure XX), and
obtained analogous results. With 216 possible agreement patterns, 159
were realized in the data, and our proposed method became much faster,
finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
relatively unchanged from the first implementation. This demonstrates
the way that the computational complexity of our method depends on the
number of unique agreement patterns, and how significant computional
gains can be made by simplifying the construction of the comparison
vectors. We also note that estimates of the \(\mathbf{m}\) and
\(\mathbf{u}\) paramters under each method are very similar, as shown in
Figure XX.

\hypertarget{violations-of-one-to-one-matching}{%
\subsection{Violations of One-to-One
Matching}\label{violations-of-one-to-one-matching}}

Figure XX shows the posterior distribution for \(D\), the number of
duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
see that \texttt{fabl} consistently overmatches within each Gibbs
iteration when compared to \texttt{BRL}, which is to be expected because
\texttt{BRL} explicitly prevents matches that violate one-to-one
matching throughout the entire sampler. Most of such matchings are due
to the randomness in the sampling procedure, and they occur sporadically
throughout the sampler in such a way that does not measurably influence
the eventual Bayes estimate.

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}

@@

\caption{Posterior distribution of overlap across the two files. The solid lines show the Bayes estimate for the amonut of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:unnamed-chunk-5}
\end{figure}

@@

\begin{figure}

{\centering \includegraphics[width=6.51in]{../notes/figures/el_salvador/rosa_records}

@@

\caption{Example of linkage situation with multiple plausible matches}\label{fig:unnamed-chunk-6}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=22.86in]{../notes/figures/el_salvador/bad_mixing}

@@

\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:unnamed-chunk-7}
\end{figure}

The records shown in Figure XX provide one such instance. Note these
records present a near violation of our assumption that there are no
duplications within files; we continue to assume that records 825 and
826 in ERTL correspond to different individuals, but their records are
nearly identical. Additionally, using the modified Levenstein distance
from Sadinle 2017, the comparison vectors \(\gamma_{2776, 825}\) and
\(\gamma_{2776, 826}\) are exactly identical.

Figure XX shows the values that \(Z_{825}\) and \(Z_{826}\) take on
throughout the Gibbs sampler, and demonstrates how each method handles
this situation. Under \texttt{fabl}, both records in \(B\) match to the
same record in \(A\) throughout the Gibbs process, creating consistent
violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
creates one matching configuration stays there for a while. However, if
one pair ``unmatches,'' then the other record has a chance to latch on.
Then, the Gibbs process is stuck with that matching status for a while,
resulting in a Gibbs process with poor mixing. Additionally,
\texttt{fabl} allows the modeller to inspect records with mulitple
plausible matches, and if they desire, to then choose the record pairing
with the highest posterior probability. \texttt{BRL} in contrast, in
strictly enforcing one-to-one matching throughout the sampler, can lead
to situations where none of the plausible matches reach the threshold to
be identified through the Bayes estimate.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

>>>>

\documentclass[12pt,letterpaper]{article}

\usepackage[hypertexnames=false]{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}

% Custom packages
\usepackage{amsthm, amssymb}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{tabularx}
\usepackage{tikz}           % vector graphics
\usepackage{bm}             % bold math symbols
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{import}
\usepackage{titling}
\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
  calc, shadows, plotmarks}

\graphicspath{{./notes/figures/}}

% Custom commands
\let\oldvec\vec
\renewcommand\vec{\bm}
\newcommand{\simfn}{\mathtt{sim}} % similarity function
\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
\newcommand{\distfn}{\mathtt{dist}} % distance function
\newcommand{\valset}{\mathcal{V}} % attribute value set
\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
\newcommand{\euler}{\mathrm{e}} % Euler's constant
\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
\def\spacingset#1{\renewcommand{\baselinestretch}%
  {#1}\small\normalsize} \spacingset{1}

\newtheorem*{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem*{definition}{Definition}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\sectionfont{\large\nohang\centering\MakeUppercase}

\title{Efficient and Scalable Entity Resolution with Fast Beta Linkage  (fabl)}
\author{Brian Kundinger\textsuperscript{a} \and
  Jerome Reiter\textsuperscript{a} \and 
  Rebecca C.~Steorts\textsuperscript{b}}
\date{
  \textsuperscript{a}Department of Statistical Science, Duke University\\
  \textsuperscript{b}Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\
  \today}

\begin{document}
\maketitle

\bigskip
\begin{abstract}
Abstract
\end{abstract}

\noindent%
{\it Keywords:} record linkage, Bayesian methods, paralellized computing, streaming algorithms

\newpage
\spacingset{1.5}

\section{Introduction}
\label{sec:introduction}

Entity resolution (or record linkage) is the task of identifying
duplicate records across multiple data sources. In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years. In others, we seek to perform some other statistical procedure, such as survival analysis or regression, using linked datasets, and one must  propagate linkage uncertainty into the subsequent analysis.

Record linkage is straightforward when provided unique identifiers or highly reliable
information, but becomes difficult when data sources are plagued with
error. This motivates the use of probabilistic record linkage techniques that are able to model the reliability of the information and which are more robust to errors than are deterministic methods. Unfortunately, probabilistic record linkage is computationally intense, and many entity resolution tasks are large in nature,
motivating the search for scalable and efficient record linkage techniques.

In this paper, we propose an extention to the \emph{Beta Record Linkage}
(\texttt{BRL}) method proposed proposed by \citep{sadinle2017} for this very end.  We increase the speed of the \texttt{BRL} procedure through the modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity. Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks. To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable'')

In what follows, Section 2 describes previous work in record linkage that lays the foundation for our method, Section
3 provides our model and the Gibbs sampler used for posterior inference,
and Section 4 describes strategies used for more efficient and scalable
implementation. Then, Section 5 demonstrates the accuracy and speed of
out method through two simulation studies, and Section 6 reveals
interesting properties of the method through a case study of El
Salvadoran casualty records. Lastly, our discussion highlights open
questions in the record linkage literature and motivates future work.

\section{Related work}
\label{sec:related}

\paragraph{Fellegi and Sunter Method}
Most record linkage techniques are derived from the seminal 1969 paper
by Fellegi and Sunter: ``A Theory for Record Linkage''. The defining
characteristic of their approach was to transform the sets of records,
which often contain text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates an
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
score on the \(f^{th}\) field. For ease of modeling and computation, we
restrict these similarity scores to be discrete, ordinal variables, and
the construction of these is left to the modeler. We adopt the
convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
is common to use use binary 1-2 variables to indicate exact matching,
and 1-2-3 variables to provide an option for partial matching. For text
data, we calculate similarity based on Levenstein distance or some other
text similarity score, and bin these scores to integers for use in the
model. Jaro (1989) provided guidelines for appropriate binning for the
comparison vectors, but details of implementation are generally context
specific.

@@

\paragraph{Sadinle's Method for Bipartite Matching} \texttt{BRL} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have
already been matched, and then samples a potential link. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

\paragraph{Deterministic Blocking}

\paragraph{Other Approach (Will probably remove)}Outside of the Fellegi-Sunter framework, some researchers have developed
methods that model the data directly, not a derived set of comparison
vectors. As an early example, Steorts et al 2016 produced
\texttt{blink}, which had the advantage of a computational complexity
that grew linearly with the data (as opposed to quadratically within
Fellegi Sunter), but was plagued with slow mixing times and was unable
to incorporate text data. Later, Steorts et al 2020 improved on this
method with \texttt{dblink}, which was able to simulate text data within
its Gibbs sampler by drawing from an empirical distribution, and
remedied its slow mixing through a probabilistic blocking approach.
However, this method remains computationally intensive and requires
distributed computing to manage even moderate linkage problems in
acceptable time. Because these methods fall outside of the
Fellegi-Sunter framework, we do not consider them further in this paper.

\section{A scalable model for Bayesian Record Linkage}
\label{sec:model}

@@

\hypertarget{model-specification}{%
	\subsection{Model Specification}
	\label{model-specification}}

@@

\hypertarget{posterior-sampling}{%
	\subsection{Posterior Sampling}
	\label{posterior-sampling}}
We work with the following factorization of the joint posterior distribution:

\[p(\mathbf{Z}, \mathbf{m}, \mathbf{u}, \lambda|\Gamma) \propto p(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) p(\mathbf{Z} | \lambda) p(\mathbf{m}, \mathbf{u}) p(\lambda)\]

@@

\hypertarget{bayes-estimate}{%
	\subsection{Bayes Estimate}
	\label{bayes-estimate}}

@@

\section{Efficient and Scalable Implementation}
\label{sec:efficiency}
Broadly speaking, we increase our computational efficiency by
recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. Let
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P = |\mathcal{H}|\) denote the total number of unique agreement
patterns. Note that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
that this bound does not scale with \(n_A\) or \(n_B\). Prior to
processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
and enumerate them \(h_1, \ldots, h_P\). When the \((i,j)\) pair
exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
Wherever possible, we conduct calculations over these \(P\) agreement
patterns rather than the \(n_A \times n_B\) record pairs.

@@

\section{Efficient Posterior Inference}\label{sec:efficient-posterior}

@@

\section{Simulation Studies}
\label{sec:simulations}

@@

\begin{figure}[ht]

	{\centering

		\includegraphics{../notes/figures/sadinle_sim_plot} 
	}

	\caption{Posterior means and credible intervals for accuracy metrics under the replication of simulation study from Sadinle 2017. For each level of overlap and each level of error, we have 100 paired sets of 500 records.}\label{fig:sadinle_simulation}

\end{figure}

@@

\begin{figure}[ht]

	{\centering \includegraphics{../notes/figures/sadinle_speed_plot2}

@@

	\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$}\label{fig:speed1}
\end{figure}

@@

\begin{figure}[t]

	{\centering \includegraphics{../notes/figures/speed_plot_fixed_nB}

@@

	\caption{Runtime for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500}\label{fig:speed2}
\end{figure}

@@

\section{Case Studies}
\label{case-studies}

We now demonstrate the power of our method through two case studies. The first is a study of documented identifiable deaths (DID) from the El Salvadoran Civil War. Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be acheived by simplifying the construction of the comparison vectors. Secondly, the case study reveals the impact of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}. The second case study is the National Long Term Care Survey (NLTCS), a much larger linkage tasks which \texttt{BRL} is unable to complete but which we complete with \texttt{fabl} with relative ease.

\paragraph{El Salvadoran Civil War}The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualities, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casuality lists. To obtain a more accurate estimate of the
casualties then, we follow Sadinle's 2017 paper and link records of El
Salvadoran civilian casualties from two sources: El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

@@

\begin{figure}[t]

	{\centering \includegraphics{../notes/figures/el_salvador/m_posterior_smallP}

@@

	\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
\end{figure}

To mirror the original implementation, we constructed the comparison
vectors using 4 levels of agreement for each field, according to the
thresholds provided in Figure XX. This took 422 seconds for our proposed
method and 240 for \texttt{BRL}. However, we observed that posterior
distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
parameters were nearly identical, and that of the
\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
realized in the data. This leads us to believe that such high number of
aggreement levels creates unnecesary distictions in the data and makes
the comparison vectors less interpretable. Therefore we re-ran our
analysis with fewer agreement levels for each field (see Figure XX), and
obtained analogous results. With 216 possible agreement patterns, 159
were realized in the data, and our proposed method became much faster,
finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
relatively unchanged from the first implementation. This demonstrates
the way that the computational complexity of our method depends on the
number of unique agreement patterns, and how significant computional
gains can be made by simplifying the construction of the comparison
vectors. We also note that estimates of the \(\mathbf{m}\) and
\(\mathbf{u}\) paramters under each method are very similar, as shown in
Figure~\ref{fig:m-and-u}.

\subsection{Violations of One-to-One Matching}
\label{violations-of-one-to-one-matching}

Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
see that \texttt{fabl} consistently overmatches within each Gibbs
iteration when compared to \texttt{BRL}, which is to be expected because
\texttt{BRL} explicitly prevents matches that violate one-to-one
matching throughout the entire sampler. Most of such matchings are due
to the randomness in the sampling procedure, and they occur sporadically
throughout the sampler in such a way that does not measurably influence
the eventual Bayes estimate.

\begin{figure}[t]

	{\centering \includegraphics{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}

@@

	\caption{Posterior distribution of overlap across the two files. The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:overlap-plot}
\end{figure}

@@

\begin{figure}[t]

	{\centering \includegraphics{../notes/figures/el_salvador/rosa_records}

@@

	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
\end{figure}

\begin{figure}[t]

	{\centering \includegraphics{../notes/figures/el_salvador/bad_mixing}

@@

	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
\end{figure}

The records shown in Figure~\ref{fig:rosa-maria} provide one such instance. Note these
records present a near violation of our assumption that there are no
duplications within files; we continue to assume that records 825 and
826 in ERTL correspond to different individuals, but their records are
nearly identical. Additionally, using the modified Levenstein distance
from Sadinle 2017, the comparison vectors \(\gamma_{2776, 825}\) and
\(\gamma_{2776, 826}\) are exactly identical.

Figure XX shows the values that \(Z_{825}\) and \(Z_{826}\) take on
throughout the Gibbs sampler, and demonstrates how each method handles
this situation. Under \texttt{fabl}, both records in \(B\) match to the
same record in \(A\) throughout the Gibbs process, creating consistent
violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
creates one matching configuration stays there for a while. However, if
one pair ``unmatches,'' then the other record has a chance to latch on.
Then, the Gibbs process is stuck with that matching status for a while,
resulting in a Gibbs process with poor mixing. Additionally,
\texttt{fabl} allows the modeler to inspect records with multiple
plausible matches, and if they desire, to then choose the record pairing
with the highest posterior probability. \texttt{BRL} in contrast, in
strictly enforcing one-to-one matching throughout the sampler, can lead
to situations where none of the plausible matches reach the threshold to
be identified through the Bayes estimate.

\paragraph{National Long Term Care Survey} The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients. It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989. The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office. State and regional office have been anomynized, presented as numbers instead of their actual values. We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context. The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466 files$, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns. We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.

With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible. Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold. Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets. We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing. We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler.

Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computatonal complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$. However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds. Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conficts are resolved to ensure one-to-one matching.

Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking. With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the seperate blocks. This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler. However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error. In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office. We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives. Even if not, the modeller may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures. However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons. Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block. Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.

\section{Discussion}
\label{discussion}

@@

\bigskip

\begin{center}
{\large\bf ACKNOWLEDGEMENTS}
\end{center}

\bibliographystyle{jasa}
{
\footnotesize
\bibliography{references}
}
